MS_data_clean

### 2021-10-08 MS_data_modeling/MS_data_explore.Rmd
Adds specialization comparisons, bee sources table.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)

@@

@@

```{r}
MS_data_format <- MS_data %>% mutate(interestBee = str_to_title(paste(Genus, sub(".*? ", "", Species)))) %>% 
  rename(degreeMS = 'Pollen host family #', specializationMS = 'Redefined pollen diet breadth') %>%
  select(interestBee, degreeMS, specializationMS) %>%
  filter(!is.na(degreeMS)) %>% arrange(interestBee)

MS_data_format
```

@@

globee_format
```

```{r}
compare_df <- merge(globee_format, MS_data_format, by = c('interestBee'), all = T)

compare_df <- compare_df %>% merge(fowler %>% rename(specialistFOWLER = specialist) %>% 
                                     select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% 
  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(degreeMS), 0, 1), 
         existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
         existsALL = existsGLOBEE*existsMS*existsFOWLER)

compare_df %>% filter(existsALL == 1)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lifecycle)
library(rvest)
library(dplyr)
library(tidyr)
library(plotly)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)
library(kableExtra)

@@

globee_5cut <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteC5') %>%
  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)

globee_T <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteW') %>%
  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)

@@

```{r}
MS_data_format <- MS_data %>% mutate(interestBee = str_to_title(paste(Genus, sub(".*? ", "", Species)))) %>% 
  rename(degreeMS = 'Pollen host family #', specializationMS = 'Redefined pollen diet breadth') %>%
  select(interestBee, degreeMS, specializationMS) %>%
  filter(!is.na(degreeMS)) %>% arrange(interestBee)
```

@@

globee_5cut_format <- globee_5cut %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE_C5 = degree) %>% 
  select(interestBee, degreeGLOBEE_C5) %>% unique() %>%
  arrange(interestBee)

globee_T_format <- globee_T %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE_T = degree) %>% 
  select(interestBee, degreeGLOBEE_T) %>% unique() %>%
  arrange(interestBee)
```

```{r}
compare_df <- merge(globee_format, MS_data_format, by = c('interestBee'), all = T)  
compare_df <- merge(compare_df, globee_5cut_format, by = c('interestBee'), all = T) 
compare_df <- merge(compare_df, globee_T_format, by = c('interestBee'), all = T)

compare_df <- compare_df %>% merge(fowler %>% rename(specialistFOWLER = specialist) %>% 
                                     select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% unique() %>%
  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(degreeMS), 0, 1), 
         existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
         existsALL = existsGLOBEE*existsMS*existsFOWLER)
```

### Add individual citation cuts in addition to raw GLOBI
### Figure out what we want to define as 'specialist' MS data

```{r}
only_globi <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 0, existsFOWLER == 0) %>% nrow()
globi_fowler <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 0, existsFOWLER == 1) %>% nrow()
globi_MS <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 1, existsFOWLER == 0) %>% nrow()
MS_fowler <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 1, existsFOWLER == 1) %>% nrow()
fowler_only <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 0, existsFOWLER == 1) %>% nrow()
MS_only <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 1, existsFOWLER == 0) %>% nrow()
all_sources <- compare_df %>% filter(existsALL == 1) %>% nrow()

counts <- c(only_globi,fowler_only,MS_only, globi_MS, globi_fowler, MS_fowler, all_sources, nrow(compare_df))
data_sources <- c('Only in GLOBI', 'Only in Fowler', 'Only in MS', 'GLOBI and MS', 'GLOBI and Fowler', 'MS and Fowler', 'All Sources', 'Total')

data.frame(data_sources, counts) %>% mutate('Percent of Data' = counts / nrow(compare_df)) %>% 
  rename('Data Source' = data_sources, 'Unique Genus Count' = counts) %>% kbl(caption = "Bee Data Sources Summary") %>%
  kable_styling()
```

```{r}
compare_df %>% filter(existsALL == 1) %>% group_by(specializationMS) %>% summarize('Mean GLOBI Degree'  = mean(degreeGLOBEE),
                                                                                   'GLOBI Degree IQR' = paste('[', round(quantile(degreeGLOBEE, 0.25),
                                                                                                                         2), 
                                                                                                                  ',', round(quantile(degreeGLOBEE, 0.75), 2),
                                                                                                            ']'),
                                                                                   'Mean GLOBI C5 Degree' = mean(degreeGLOBEE_C5),
                                                                                   'GLOBI C5 Degree IQR' = paste('[', round(quantile(degreeGLOBEE_C5, 0.25), 2),
                                                                                                                  ',', round(quantile(degreeGLOBEE_C5, 0.75), 2),
                                                                                                            ']'),
                                                                                   'Mean GLOBI T Degree' = mean(degreeGLOBEE_T),
                                                                                   'GLOBI T Degree IQR' = paste('[', round(quantile(degreeGLOBEE_T, 0.25), 2),
                                                                                                                  ',', round(quantile(degreeGLOBEE_T, 0.75), 2),
                                                                                                            ']'),
                                                                                   'Mean MS Degree' = mean(degreeMS),
                                                                                   'MS Degree IQR' = paste('[', round(quantile(degreeMS, 0.25), 2), 
                                                                                                                  ',', round(quantile(degreeMS, 0.75), 2),
                                                                                                            ']'),
                                                                                   'Percent Fowler Specialist' = mean(specialistFOWLER)
) %>% kbl(caption = 'Plant Family Degrees by MS Specialization') %>%
  kable_styling()
```

```{r}
# Bees fowler specialists but not MS specialists

compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5)
```

```{r}
# Bees NOT fowler specialists but MS specialists

compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5)
```

## Plan: want to add pollen hosts so I can filter on pollen family. Discuss how to parse the authors out of appropriate MS column.
